### 2013-03-31 parts/appendix.tex
create appendix

\appendix

%\section{Demonstration of \pref{th:approxinf}}

\section{Demonstration of \towrite{équivalence PHP / réseaux discrets}}

\todo{Preuve}