### 2014-03-18 parts/models.tex
finish proof for flattening; minor review/additions left

\section{\todo{Title?} Equivalence with other frameworks}

\subsection{Asynchronous Discrete Networks}
\todo{Detail the equivalence with ADNs (cf. the annex)}

\subsection{PH with multiple classes of priorities}
\todo{Glue up all this section}

\newcommand{\ov}{\overline}
\newcommand{\os}{\ov{s}}
\newcommand{\oPH}{\ov{\PH}}

\newcommand{\F}{F}
\newcommand{\Fopsymbol}{\F}
\newcommand{\Fop}[1]{\Fopsymbol(#1)}
\newcommand{\Fsem}[2]{\left[#1\right](#2)}
%\newcommand{\Fall}{\F}
\newcommand{\mysigma}{\sigma}

Consider a PH with $k \in \sN^*$ classes of priories:
$\ov{\PH} = (\ov{\PHs}; \ov{\PHl}; \ov{\PHa}^{\angles{k}})$.
We want to translate it to a PH with $2$ classes of priorities
$\PH = (\PHs; \PHl; \PHa^{\angles{2}})$
that respects all Criteria given in \pref{ssec:hypothesis}.

\newcommand{\PHsubl}[1][\ov{\PHl}]{#1^\diamond}

\todo{Rather define a sub-state on a subset of sorts}
\begin{definition}[Sub-states ($\PHsubl$)]
  A sub-state of $\oPH$ is an element of:
    \[\PHsubl \DEF \{ \mysigma \in \bigtimes{a \in S} \ov{\PHl}_a \mid S \subset \ov{\PHs} \} \enspace.\]
  Furthermore, if $\mysigma \in \PHsubl$ and $s \in \ov{\PHl}$, we note:
    \[\mysigma \subseteq s \EQDEF \forall a_i \in \Proc, a_i \in \mysigma \Rightarrow a_i \in s \enspace.\]
\end{definition}

We note that a state is \textit{a fortiori} a sub-state: $\ov{\PHl} \subset \PHsubl$.

A playability property (PP) as defined in \pref{def:pp} is a Boolean formula where the atoms are processes of $\ov{\PH}$.

\begin{definition}[Playability property language ($\F$)]
  \label{def:pp}
  A \emph{playability property} (PP) is an element of the language $\F$ inductively defined by:
  \begin{itemize}
    \item $\top$ and $\bot$ belong to $\F$;
    \item if $a \in \ov{\PHs}$ and $a_i \in \ov{\PHl}_a$, then $a_i \in \F$ and we call it an \emph{atom};
    \item if $P \in \F$ and $Q \in \F$, then $\neg P \in \F$, $P \wedge Q \in \F$ and $P \vee Q \in \F$.
  \end{itemize}
  If $P \in \F$ is a PP and $\mysigma \in \PHsubl$ is a sub-state,
  we note $\Fsem{P}{\mysigma}$ the \emph{evaluation} of $P$ in $\mysigma$:
%  We note $\Fsem{P}{s}$ the \emph{evaluation} of a PP $P \in \F$ in a state $s \in \ov{\PHl}$.
%  The semantics of this evaluation is the following for any state $s \in \ov{\PHl}$:
  \begin{itemize}
    \item if $P = a_i \in \ov{\PHs}_a$ is an atom, with $a \in \ov{\PHs}$, then $\Fsem{a_i}{\mysigma}$ is true iff $a_i \in \mysigma$;
    \item if $P$ is not an atom, then $\Fsem{P}{\mysigma}$ is true iff all its subformulas are inductively true in $\mysigma$
      with the classical semantics for the logic operators $\top$, $\bot$, $\neg$, $\wedge$ and $\vee$.
  \end{itemize}
\end{definition}

Because we only use classical logic operators, the formulas of Boolean logic on distribution and negation (de Morgan's laws) can be used.
We also have the following property for the negation of an atom:
\[\forall a \in \ov{\PHs}, \forall a_i \in \ov{\PHl}_a, \forall \mysigma \in \PHsubl,
  \Fsem{\neg a_i}{s} \Leftrightarrow \Fsem{\bigvee_{\substack{a_j \in \PHl_a\\a_j \neq a_i}} a_j}{s}\]
Indeed, if a process is not active in a state, this means that another process of the same sort is active.

In \pref{def:fop}, we finally define the the operator $\Fopsymbol$ which characterizes the playability of an action
given the semantics of PH (\pref{def:play}).
\begin{definition}[Playability property of an action ($\Fopsymbol : \PHh \rightarrow \F$)]\label{def:fop}
  $$\forall h \in \ov{\PHh}, \Fop{h} \equiv \hitter{h} \wedge
    \left( \bigwedge_{\substack{g \in \PHh^{(n)}\\n < \prio(h)}}
    \neg \left( \target{g} \wedge \hitter{g}\right) \right)$$
\end{definition}
%
\begin{theorem}
\label{th:ppplay}
  $h \in \ov{\PHh}$ is playable in $s \in \ov{\PHl}$ if and only if $\Fsem{\Fop{h}}{s}$.
\end{theorem}
%
\begin{proof}
  By definition of the semantics of a PH given in \pref{def:play}.
\end{proof}

\newcommand{\n}[1][h]{n^{#1}}
\newcommand{\m}[1][h,i]{m^{#1}}

Because we only use classical logic operators, we can compute the Disjunctive Normal Form (DNF) of any PP.
For any action $h \in \ov{\PHh}$, this DNF takes the form:
\[\Fop{h} \equiv \bigvee_{i \in \segm{1}{\n}} \left( \bigwedge_{j \in \segm{1}{\m}} p_{i,j} \right)\]
where $\n \in \sN$ and $\forall i \in \segm{1}{\n}, \m \in \sN^*$.
If $\n = 0$, then $\Fop{h} \equiv \bot$; this means that $h$ can never be played
due to preemptions by other actions with higher priorities.
If $\Fop{h} \not\equiv \bot$, on the other hand, then in this case $\Fop{h}$
can be seen as a disjunction of $\n$ smaller playability properties consisting only of conjunctions of atoms.
These $\n$ conjunctions can be translated to as many prioritized cooperative sorts,
thus creating a new PH model which is entailed in the restrictions depicted in \pref{ssec:hypothesis}.
%Indeed, for any $i \in \segm{1}{\n}$, the PP $\bigwedge_{j \in \segm{1}{\m}} p_{i,j}$
In this second case, we denote, for any $i \in \segm{1}{\n}$:
$\PHdep{i}{h} = \{ \PHsort(p_{i,j}) \mid j \in \segm{1}{\m} \}$.

\todo{Glue up}
\begin{theorem}
\label{th:ppplaysubset}
  $h \in \ov{\PHh}$ is playable in $s \in \ov{\PHl}$ if and only if
  $\exists \mysigma \subseteq s, \Fsem{\F(h)}{\mysigma}$.
%  \[\exists i \in \segm{1}{\n}, \exists \mysigma \in \bigtimes{b \in \PHdep{i}{h}} \PHl_b,
%    \mysigma \subseteq s \wedge \Fsem{\F(h)}{\mysigma} \enspace.\]
\end{theorem}
%
\begin{proof}
  ($\Rightarrow$) If $\Fsem{\F(h)}{s}$ then $\F(h) \not\equiv \bot$ and, by property of a DNF,
    at least one of the $\n$ conjunctions of $\F(h)$ is true in $s$.
    Suppose the $i$\textsuperscript{th} conjunction is true in $s$, with $i \in \segm{1}{\n}$;
    then we have: $\forall j \in \segm{1}{\m}, p_{i,j} \in s$.
    Let $\mysigma \in \bigtimes{b \in \PHdep{i}{h}} \PHl_b$ with $\forall b \in \PHdep{i}{h}, \PHget{\mysigma}{b} = \PHget{s}{b}$.
    We immediately have: $\mysigma \subseteq s$,
    and, by construction of $\PHdep{i}{h}$, $\Fsem{\F(h)}{\mysigma}$.

  ($\Leftarrow$) Trivial.
\end{proof}

\newcommand{\flats}[1]{\lfloor #1 \rfloor}
\newcommand{\unflats}[1]{\lceil #1 \rceil}

\begin{definition}
  \label{def:flattening}
  If $k \in \sN^*$, we note $\PHflat(\ov{\PHs}; \ov{\PHl}; \ov{\PHa}^{\angles{k}}) = (\PHs; \PHl; \PHa^{\angles{2}})$
  the flattening of the PH with $k$ classes of priorities $\ov{\PHs}; \ov{\PHl}; \ov{\PHa}^{\angles{k}}$, with:
  \begin{itemize}
    \item $\PHs = \ov{\PHs} \cup \PHs_f$
      where $\PHs_f = \{ f^{h,i} \mid h \in \ov{\PHh} \wedge \n \geq 1 \wedge i \in \segm{1}{\n} \}$;
    \item $\PHl = \left( \underset{a \in \ov{\PHs}}{\times} \PHl_{a} \right) \times
      \left(\underset{f^{h,i} \in \PHs_f}{\times} \PHl_{f^{h,i}} \right)$,
      where $\forall a \in \ov{\PHs}, \PHl_{a} = \ov{\PHl}_{a}$, and
      $\forall f^{h,i} \in \PHs_f,
      \PHl_{f^{h,i}} = \{ f^{h,i}_\mysigma \mid \mysigma \in \underset{b \in \PHdep{i}{h}}{\times} \PHl_{b} \}$;
    \item $\PHh^{(1)} = \{ \PHhit{a_k}{f^{h,i}_\mysigma}{f^{h,i}_{\mysigma'}} \mid
      h \in \ov{\PHh} \wedge
      %\n \geq 1 \wedge i \in \segm{1}{\n} \wedge
      f^{h,i} \in \PHs_f \wedge
      a \in \PHdep{i}{h} \wedge a_k \in \PHl_a \wedge
      f^{h,i}_\mysigma , f^{h,i}_{\mysigma'} \in \PHl_{f^{h,i}} \wedge
      \PHget{\mysigma}{a} \neq a_k \wedge \mysigma' = \mysigma \Cap \{ a_k \} \}$;
%      \todo{From here: ...} \PHget{\mysigma'}{a} = a_k \wedge
%      (\forall b \in \PHdep{i}{h} \setminus \{ a \}, \PHget{\mysigma'}{b} = \PHget{\mysigma}{b}) \}$;
%      \todo{... can be replaced by: $\mysigma' = \mysigma \Cap \{ a_k \}$}
    \item $\PHh^{(2)}=\{ \PHhit{f^{h,i}_\mysigma}{\target{h}}{\bounce{h}} \mid
      h \in \ov{\PHh} \wedge 
      f^{h,i} \in \PHs_f \wedge
      f^{h,i}_\mysigma \in \PHl_{f^{h,i}} \wedge \Fsem{\F(h)}{\mysigma} \}$.
  \end{itemize}
  Given a state $s \in \PHl$, $\unflats{s} = \os$ is the corresponding state in $\ov{\PHl}$:
  $\forall a \in \ov{\PHs}, \PHget{\os}{a} = \PHget{s}{a}$.

  \noindent
  Given a state $\os \in \ov{\PHl}$, $\flats{\os} = s$ is the corresponding state in $\PHl$:
  $\forall a \in \ov{\PHs}, \PHget{s}{a} = \PHget{\os}{a}$
  and $\forall f^{h,i} \in \PHs_f, \PHget{s}{f^{h,i}} = f^{h,i}_\mysigma$ with $f^{h,i}_\mysigma \in \PHl_{f^{h,i}}$
  and $\forall b \in \PHdep{i}{h}, \PHget{\mysigma}{b} = \PHget{\os}{b}$.
\end{definition}

\begin{theorem}[$(\ov{\PHs}; \ov{\PHl}; \ov{\PHa}^{\angles{k}}) \approx \PHflat(\ov{\PHs}; \ov{\PHl}; \ov{\PHa}^{\angles{k}})$]
\label{thm:bisimPHP}
  Let $\ov{\PH} = (\ov{\PHs}; \ov{\PHl}; \ov{\PHa}^{\angles{k}})$ a PH with $k$ classes of priorities,
  and $\PH = \PHflat(\ov{\PHs}; \ov{\PHl}; \ov{\PHa}^{\angles{k}}) = (\PHs; \PHl; \PHa^{\angles{2}})$ its flattening.
  \begin{enumerate}
    \item \label{php2ph} $\forall \os, \os' \in \ov{\PHl}$,
      $\os \PHPtrans[\oPH] \os' \Longrightarrow \flats{\os} \PHPtrans[\oPH]^* \flats{\os'}$,
      where $\PHPtrans[\oPH]^*$ is a finite sequence of $\PHPtrans[\oPH]$ transitions.
    \item \label{ph2php} $\forall s, s' \in \PHl$,
      $s \PHPtrans s' \Longrightarrow \unflats{s} = \unflats{s'} \vee
      \unflats{s} \PHPtrans \unflats{s'} \enspace.$
  \end{enumerate}
\end{theorem}
%
\begin{proof}
  (\ref{php2ph}) Let $\flats{\os} = s$.
    Given the dynamics of a PH (\pref{def:play}), if $\os \PHPtrans[\PH] \os'$,
    then there exists $h \in \ov{\PHh}$ playable in $\os$;
    therefore, $\os' = \os \PHplay h$ and $\target{h} \in \os$.
    Furthermore, given \pref{th:ppplay}, $\Fsem{\F(h)}{\os}$,
    and given \pref{th:ppplaysubset}, there exists $\mysigma \subseteq s$ so that $\Fsem{\F(h)}{\mysigma}$.
    Therefore, by construction of $\PH$ (\pref{def:flattening}) there exists
    $g = \PHhit{f^{h,i}_\mysigma}{\target{h}}{\bounce{h}} \in \PHh^{(2)}$.
    By construction of $s$ (\pref{def:flattening}), $\PHget{s}{f^{h,i}} = f^{h,i}_\mysigma$ and $\target{h} = \target{g} \in s$.
    Therefore, $g$ is playable in $s$.
    In $s \PHplay g$, the only playable actions are those in $\PHh^{(1)}$ having $\bounce{h} = \bounce{g}$ as hitter
    and updating cooperative sorts, allowing to reach $\flats{\os'}$ in a finite number of actions.
    Thus, $\flats{\os} \PHPtrans[\oPH]^* \flats{\os'}$.

  (\ref{ph2php}) Let $\os = \unflats{s}$.
    Given the dynamics of a PH (\pref{def:play}), if $s \PHPtrans s'$,
    then there exists $g \in \PHh$ playable in $s$; therefore, $s' = s \PHplay g$ and $\target{g} \in s$.
    If $\prio(g) = 1$ then only the active process of a cooperative sort has evolved, and $\unflats{s} = \unflats{s'}$.
    Otherwise, $\prio(g) = 2$; we note in this case: $g = \PHhit{f^{h,i}_\mysigma}{b_j}{b_k}$.
    By construction of the flattening (\pref{def:flattening}), there exists $h \in \ov{\PHh}$ so that
    $\target{h} = b_j$, $\bounce{h} = b_k$ and $\Fsem{\F(h)}{\mysigma}$.
    If $g$ is playable, this means that no other action in $\PHh^{(1)}$ is playable, and especially the cooperative sort
    $f^{h,i}$ is already updated; therefore, $\mysigma \subseteq s$.
    By \pref{th:ppplaysubset}, it comes that $h$ is playable in $\os$.
    Thus, $\unflats{s} \PHPtrans \unflats{s'}$.
\end{proof}