### 2014-03-19 parts/appendix.tex
several fixes + refactoring of files

\appendix

% Fix appendix references
\renewcommand*{\thesection}{\Alph{section}}

% Proof of under-approximation
\input{parts/suppl-sa.tex}

% Proofs for flattening
\input{parts/suppl-ph-php.tex}

% Proof of ADN equivalence
\input{parts/suppl-ph-dn.tex}