### 2014-03-19 parts/models-ph-dn.tex

\subsection{Asynchronous Discrete Networks}
\todo{Detail the equivalence with ADNs (cf. the annex)}