### 2014-03-19 parts/models.tex

\section{\todo{Title?} Equivalence with other frameworks}

% Equivalence with ADNs
\input{parts/models-ph-php.tex}

% Flattening of a PH with k priorities
\input{parts/models-ph-dn.tex}