### 2014-03-19 parts/suppl-ph-php.tex

\section{Proofs about flattening (\pref{ssec:flattening})}
\label{suppl:demoflattening}