### 2014-03-20 parts/models.tex
simplifications in flattening + begin example

\section{Encodings in Process Hitting with Prioritized Cooperations}

% Equivalence with ADNs
\input{parts/models-ph-dn.tex}

% Flattening of a PH with k priorities
\input{parts/models-ph-php.tex}