### 2014-04-14 parts/models.tex
typo: "Prioritized" -> "Prioritised"

\section{Encodings in Process Hitting with Prioritised Cooperations}
\label{sec:encodings}

% Equivalence with ADNs
\input{parts/models-ph-dn.tex}

% Flattening of a PH with k priorities
\input{parts/models-ph-php.tex}