### 2014-12-22 main.tex
begin transformation PH to AAN

\documentclass{elsarticle}

\usepackage[UKenglish]{babel}
\usepackage[utf8]{inputenc}
\usepackage[T1]{fontenc}

\usepackage{amsmath}  % Maths
\usepackage{amsfonts} % Maths
\usepackage{amssymb}  % Maths
\usepackage{stmaryrd} % Maths (crochets doubles)

\usepackage{url}     % Mise en forme + liens pour URLs
\usepackage{hyperref}
\usepackage{array}   % Tableaux évolués

\usepackage{comment}

% Theorems and definitions
\newdefinition{definition}{Definition}
\newdefinition{condition}{Condition}
\newtheorem{theorem}{Theorem}
\newtheorem{lemma}{Lemma}
\newproof{example}{Example}
\newproof{rawproof}{Proof}

\newenvironment{proof}{\begin{rawproof}}{\hfill$\Box$\end{rawproof}}

% Pretty references
\usepackage{prettyref}
\newrefformat{def}{Def.~\ref{#1}}
\newrefformat{fig}{Fig.~\ref{#1}}
\newrefformat{tab}{Table~\ref{#1}}
\newrefformat{lem}{Lemma~\ref{#1}}
\newrefformat{th}{Theorem~\ref{#1}}
\newrefformat{co}{Corollary~\ref{#1}}
\newrefformat{sec}{Sect.~\ref{#1}}
\newrefformat{ssec}{Subsect.~\ref{#1}}
\newrefformat{suppl}{Appendix~\ref{#1}}
\newrefformat{cr}{Condition~\ref{#1}}
\newrefformat{eq}{Eq.~\eqref{#1}}
\def\pref{\prettyref}

\usepackage{tikz}
\newdimen\pgfex
\newdimen\pgfem
\usetikzlibrary{arrows,shapes,shadows,scopes}
\usetikzlibrary{positioning}
\usetikzlibrary{matrix}
\usetikzlibrary{decorations.text}
\usetikzlibrary{decorations.pathmorphing}

\input{macros/macros}
\input{macros/macros-ph}
\input{macros/macros-abstr}

% Macros
\newcommand{\sN}{\mathbb{N}}
\newcommand{\sNN}{\mathbb{N}^\bullet}
\def\DEF{\stackrel{\Delta}=}
\def\EQDEF{\stackrel{\Delta}\Leftrightarrow}
\newcommand{\segm}[2]{\llbracket #1; #2 \rrbracket}
\def\indexes#1{\mathbb{I}^{#1}}
\def\f#1{\mathsf{#1}}
\def\prio{\mathsf{prio}}
\newcommand{\bottom}{\perp}
\newcommand{\stable}{\mathsf{stable}}
\newcommand{\update}{\mathsf{update}}
\newcommand{\components}{\PHs}
%\newcommand{\cs}{CS}
\newcommand{\cs}{\textcolor{red}{\Delta}}

% Restriction notation
\def\restriction#1#2{\mathchoice
              {\setbox1\hbox{${\displaystyle #1}_{\scriptstyle #2}$}
              \restrictionaux{#1}{#2}}
              {\setbox1\hbox{${\textstyle #1}_{\scriptstyle #2}$}
              \restrictionaux{#1}{#2}}
              {\setbox1\hbox{${\scriptstyle #1}_{\scriptscriptstyle #2}$}
              \restrictionaux{#1}{#2}}
              {\setbox1\hbox{${\scriptscriptstyle #1}_{\scriptscriptstyle #2}$}
              \restrictionaux{#1}{#2}}}
\def\restrictionaux#1#2{{#1\,\smash{\vrule height .8\ht1 depth .85\dp1}}_{\,#2}}

%\renewcommand{\restriction}[2]{#1_{#2}}

% Commandes À FAIRE
\usepackage{color} % Couleurs du texte
%\definecolor{darkgreen}{rgb}{0,0.5,0}
%\newcommand{\towrite}[1]{\textcolor{darkgreen}{[#1]}}
\newcommand{\todo}[1]{\textcolor{red}{\textbf{[#1]}}}

% Ajouts et progression
\def\modMF#1{\textcolor{teal}{#1}}
\def\moda#1{\textcolor{blue}{#1}}
% \def\modLP#1{\textcolor{magenta}{#1}}
% \def\modMM#1{\textcolor{blue}{#1}}
% \def\modOR#1{\textcolor{olive}{#1}}
%\def\modMF#1{#1} \def\moda#1{#1} \def\modLP#1{#1} \def\modMM#1{#1} \def\modOR#1{#1}

% Special notations
\newcommand{\ie}{i.e.,\ }
\newcommand{\Ie}{I.e.,\ }
\newcommand{\eg}{e.g.,\ }
\newcommand{\Eg}{E.g.,\ }
\newcommand{\resp}{resp.\ }
\newcommand{\Resp}{Resp.\ }

% Césures
\hyphenation{pa-ra-me-tri-za-tion}
\hyphenation{pa-ra-me-tri-za-tions}

\def\lastname{Folschette \emph{et al.}}

\begin{document}
\begin{frontmatter}
\title{Sufficient Conditions for Reachability in Automata Networks with Priorities}

\author[irccyn]{Maxime Folschette}
\ead{Maxime.Folschette@irccyn.ec-nantes.fr}
\author[lri,amib]{Loïc Paulevé}
\author[irccyn]{Morgan Magnin}
\author[irccyn]{Olivier Roux}

\address[irccyn]{LUNAM Universit\'e, \'Ecole Centrale de Nantes, IRCCyN UMR CNRS 6597\\
(Institut de Recherche en Communications et Cybern\'etique de Nantes)\\
1 rue de la No\"e - B.P. 92101 - 44321 Nantes Cedex 3, France.}

\address[lri]{CNRS, Laboratoire de Recherche en Informatique (LRI)\\
		Université Paris-Sud - CNRS UMR 8623, France}
\address[amib]{AMIB group, Inria Saclay, France}{}

\begin{abstract}
The Process Hitting is a recently introduced framework designed for the modelling of concurrent systems.
Its originality lies in a compact representation of both components of the model and its corresponding actions:
each action can modify the status of a component, and is conditioned by the status of at most one other component.
This allowed to define very efficient static analysis based on local causality to compute reachability properties.
However, in some cases, this modelling lacks some preemption features
that would allow to model relative time constraints between interactions.

In this paper, we propose an extension of the Process Hitting framework
by using classes of priorities in order to improve its expressivity.
Furthermore, we develop a new refinement for the under-approximation of the static analysis
to give accurate results on Process Hitting models.
This method applies to a specific sub-class of models,
but this class turns out to be sufficiently expressive to represent any
Process Hitting model with an arbitrary number of classes of priorities,
thus guaranteeing a wide field of application of our method.

Finally, we also show that the priorities
allow to represent accurate Boolean gates in order to model
cooperations between components
(that is, for instance, when two components are supposed to interact with a third one
only when they are in a given configuration)
which was not possible before.
This allows to accurately model
Asynchronous Discrete Networks, either Boolean or multivalued,
and efficiently under-approximate reachability properties
in these models.
We illustrate this on the model-checking of a signalling network of 94
components, which is unprecedented.
\end{abstract}
\begin{keyword}
discrete networks \sep
abstract interpretation \sep
reachability \sep
qualitative models \sep
systems biology
\end{keyword}
\end{frontmatter}

\input{parts/intro}

\input{parts/def-ph}

\input{parts/sa}

\input{parts/models}

\input{parts/example}

\input{parts/ccl}

\section*{References}
\bibliographystyle{elsarticle-num}
\bibliography{biblio}

\input{parts/appendix}

\end{document}