### 2014-12-23 parts/appendix.tex
finish transformation on the appendix

\appendix

% Fix appendix references
\renewcommand*{\thesection}{\Alph{section}}

% Proof of under-approximation
\input{parts/suppl-sa.tex}

% % Proof of ADN equivalence
% \input{parts/suppl-ph-dn.tex}
% 
% % Proofs for flattening
% \input{parts/suppl-ph-php.tex}