### 2014-12-23 parts/models.tex
fix titles and refs

\section{Encodings in Asynchronous Automata Networks}
\label{sec:encodings}

% Equivalence with ADNs
\input{parts/models-ph-dn.tex}

% Flattening of a PH with k priorities
\input{parts/models-ph-php.tex}